---
title: "snowcrabs_transcriptome"
author: "Courtney Skalley"
date: "2023-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Download Software

```{r}
library(tidyverse)
```

Check that blast software has been downloaded

```{bash}
ls /home/shared/
```

# Make blast database

Download uniprot/swisprot fasta, rename with year and version, then unzip the file.

```{bash, blast database}
cd ../data
curl -O https://ftp.uniprot.org/pub/databases/uniprot/current_release/knowledgebase/complete/uniprot_sprot.fasta.gz
mv uniprot_sprot.fasta.gz uniprot_sprot_r2023_01.fasta.gz
gunzip -k uniprot_sprot_r2023_01.fasta.gz
ls ../data
```

```{bash}
/home/shared/ncbi-blast-2.11.0+/bin/makeblastdb \
-in ../data/uniprot_sprot_r2023_01.fasta \
-dbtype prot \
-out ../blastdb/uniprot_sprot_r2023_01
```

# Download transcriptomes

There are three available transcriptomes for opilio crabs: molting gland, eyestalk ganglia, and hepatopancreas. Here, we will blast the molting gland and eyestalk ganglia transcriptomes.

C. opilio molting gland (mg) transcriptome:

```{bash}
cd ../data
curl -o "c_opilio_mg_transcriptome.fasta" -H "Accept: text/plain; format=tsv" "https://gannet.fish.washington.edu/seashell/snaps/HBXI01.fasta"
```

C. opilio eyestalk ganglia (eg) transcriptome:

```{bash}
cd ../data
curl -o "c_opilio_eg_transcriptome.fasta" -H "Accept: text/plain; format=tsv" "https://gannet.fish.washington.edu/seashell/snaps/HBWO01.fasta"
```

### Examine fasta files

Eyestalk ganglia (eg) transcriptome:

```{bash}
head ../data/c_opilio_eg_transcriptome.fasta
echo "How many sequences are there?"
grep -c ">" ../data/c_opilio_eg_transcriptome.fasta
```

Molting gland (mg) transcriptome:

```{bash}
head ../data/c_opilio_mg_transcriptome.fasta
echo "How many sequences are there?"
grep -c ">" ../data/c_opilio_mg_transcriptome.fasta
```

# Run blast

### Blast molting gland (mg) transcriptome:

```{bash}
/home/shared/ncbi-blast-2.11.0+/bin/blastx \
-query ../data/c_opilio_mg_transcriptome.fasta \
-db ../blastdb/uniprot_sprot_r2023_01 \
-out ../output/c_opilio_mg_uniprot_blastx.tab \
-evalue 1E-20 \
-num_threads 20 \
-max_target_seqs 1 \
-outfmt 6
```

Examine blastx table for molting gland

```{bash}
head -2 ../output/c_opilio_mg_uniprot_blastx.tab
wc -l ../output/c_opilio_mg_uniprot_blastx.tab
```

### Blast eyestalk ganglia (eg) transcriptome:

```{bash}
/home/shared/ncbi-blast-2.11.0+/bin/blastx \
-query ../data/c_opilio_eg_transcriptome.fasta \
-db ../blastdb/uniprot_sprot_r2023_01 \
-out ../output/c_opilio_eg_uniprot_blastx.tab \
-evalue 1E-20 \
-num_threads 20 \
-max_target_seqs 1 \
-outfmt 6
```

Examine blastx table for eyestalk ganglia

```{bash}
head -2 ../output/c_opilio_eg_uniprot_blastx.tab
wc -l ../output/c_opilio_eg_uniprot_blastx.tab
```

# Download uniprot table

```{bash}
cd ../data
curl -o "uniprot_table_r2023_01.tab" -H "Accept: text/plain; format=tsv" "https://gannet.fish.washington.edu/seashell/snaps/uniprot_table_r2023_01.tab"
```

Examine uniprot table

```{bash}
head -2 ../data/uniprot_table_r2023_01.tab
```

# Format tables

## Separate pipes (\|) into tabs for mg

```{bash}
tr '|' '\t' < ../output/c_opilio_mg_uniprot_blastx.tab | head -2
```

```{bash}
tr '|' '\t' < ../output/c_opilio_mg_uniprot_blastx.tab \
> ../output/c_opilio_mg_uniprot_blastx_sep.tab
```

## Separate pipes (\|) into tabs for eg

```{bash}
tr '|' '\t' < ../output/c_opilio_eg_uniprot_blastx.tab | head -2
```

```{bash}
tr '|' '\t' < ../output/c_opilio_eg_uniprot_blastx.tab \
> ../output/c_opilio_eg_uniprot_blastx_sep.tab
```

Examine molting gland (mg) table

```{bash}
head -2 ../output/c_opilio_mg_uniprot_blastx_sep.tab 
```

Examine eyestalk ganglia (eg) table

```{bash}
head -2 ../output/c_opilio_eg_uniprot_blastx_sep.tab 
```

# Save blastx table as a local object
### Molting gland

```{r}
bltabl.mg <- read.csv("../output/c_opilio_mg_uniprot_blastx_sep.tab", sep = '\t', header = FALSE)
```

### Eyestalk ganglia

```{R}
bltabl.eg <- read.csv("../output/c_opilio_eg_uniprot_blastx_sep.tab", sep = '\t', header = FALSE)
```

# Save database reference table as a local object

```{r}
spgo <- read.csv("../data/uniprot_table_r2023_01.tab", sep = '\t', header = TRUE)
```

```{r}
str(spgo)
```

#Join tables using kbl 
Specify which columns to include and format names more cleanly

## Molting gland:

```{r}
left_join(bltabl.mg, spgo,  by = c("V5" = "Entry")) %>%
  write_delim("../output/blast_annot_go_mg.tab", delim = '\t')
```

## Ganglia eyestalk:

```{r}
left_join(bltabl.eg, spgo,  by = c("V5" = "Entry")) %>%
  write_delim("../output/blast_annot_go_eg.tab", delim = '\t')
```

Examine at annotated table for molting gland (mg)

```{r}
annot_tab <- read.csv("../output/blast_annot_go_mg.tab", sep = '\t', header = TRUE)
head(annot_tab)
```

Examine at annotated table for eyestalk ganglia (eg)

```{r}
annot_tab_eg <- read.csv("../output/blast_annot_go_eg.tab", sep = '\t', header = TRUE)
head(annot_tab)
```


# Annotate DGE table
Join by V3 (in annot_taba) and column with "HBXI01000004.1" (DEGtab)

Separate columns that contain "|". 

```{bash}
# Problem: DEGlist.tab is 1363 obs of 1 variable, so there aren't columns for me to separate out. 
tr '|' '\t' < ../output/DEGlist.tab \
> ../output/DEGlist_sep.tab  #rename with "_sep"
```

Compare the original table to separated table
```{bash}
head -2 ../output/DEGlist.tab
```
```{bash}
head -2 ../output/DEGlist_sep.tab #not actually separated, ENA, HBXI.. and HBXI....1 all within quotations (one column)
```
```{r}
DEGtab <- read.csv("../output/DEGlist_sep.tab", sep = ' ', header = TRUE)
# head(DEGtab) #this is empty; only 1 variable
```

```{bash}
head ../output/DEGlist_sep.tab
```

```{r}
# need to get DEGtab or annotated matrix in correct format to join columns
left_join(DEGtab, annot_tab, by = c("FIGURE THIS OUT" = "V3")) %>%
  write_delim("../output/annot_DEG.tab", delim = '\t')
```


